---
title: "assignment1"
author: "Chenfeng Hao"
date: "9/10/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.2

(a)

```{r}
x <- seq(from = 0, to = 4, by = .1)
error <- rnorm(length(x), mean = 0, sd = 3.9)
y <- 30 + 10 * x + error
plot(x, y)
```

(b)

```{r}
x <- seq(from = 0, to = 4, by = .1)
error <- rnorm(length(x), mean = 0, sd = 10)
y <- 30 + 10 * x + error
plot(x, y)
```

1.4 

(a) Generalizing from sample to population 

Predict voter shares from polls on the presidential election.  

(b) Generalizing from treatment to control  

Determine whether a vaccine is effective at preventing a certain disease from results in trials.   

(c) Generalizing from observed measurements to the underlying constructs of interests 

Determine user preferences for TV shows based on user playing history.   


1.5 

(a) Forecasting/classification 

Predict the performance of a stock based on its history and current conditions.  

(b) Exploring associations 

Determine whether a particular risk factor, say high consumption of red meat, is associated with a higher rate of cancer diagnosis.

(c) Extrapolation 

Use vaccine trial results to determine the potential effects of the vaccine on the general population. 

(d) Causal inference 

Determine whether a particular intervention is effective at improving the social and economic prospects of young children.  

1.10

TBD