---
title: "Assignment 4"
author: "Chenfeng Hao"
date: "`r format(Sys.time(), tz='EST', '%b %d, %Y')`"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


4.1 Comparison of proportions: A randomized experiment is performed within a survey. 1000  people are contacted. Half the people contacted are promised a $5 incentive to participate, and  half are not promised an incentive. The result is a 50% response rate among the treated group  and 40% response rate among the control group. Give an estimate and standard error of the average treatment effect. 

**Answer:**  
treatment effect: (0.5 * 500 - 0.4 * 500) / 1000 = `r (0.5 * 500 - 0.4 * 500) / 1000`  
standard error of the average treatment effect: $se_{treatment} = \sqrt{0.5 * 0.5 / 500}$ = `r round(sqrt(0.5 * 0.5 / 500), 2)`, $se_{control} = \sqrt{0.4 * 0.6 / 500}$ = `r round(sqrt(0.4 * 0.6 / 500), 2)`, $se_{difference} = \sqrt{se_{treatment}^2 + se_{control}^2}$ = `r round(sqrt(0.02^2 + 0.02^2),2)`

4.2 Choosing sample size: You are designing a survey to estimate the gender gap: the difference in  support for a candidate among men and women. Assuming the respondents are a simple random  sample of the voting population, how many people do you need to poll so that the standard error  is less than 5 percentage points? 

**Answer:**  
$se = \sqrt{se_{men}^2 + se_{women}^2} = \sqrt{(0.5 * 0.5) / n + (0.5 * 0.5) / n} = 0.05, n = 200$, so at least 200 people are needed.

4.3 Comparison of proportions: You want to gather data to determine which of two students is a  better basketball shooter. One of them shoots with 30% accuracy and the other is a 40% shooter.  Each student takes 20 shots and you then compare their shooting percentages. What is the  probability that the better shooter makes more shots in this small experiment? 

**Answer:**  
The simulation approach
\
```{r}
numofsims <- 100000
x1 <- rbinom(numofsims, 20, .30)
x2 <- rbinom(numofsims, 20, .40)
moreshots <- x2 > x1
mean(moreshots)
```

\
The other approach
```{r}
prob <- 0

for (i in 20:1) {
  prob <- prob + dbinom(i, 20, .4) * pbinom(i - 1, 20, .3)
}

prob
```



4.6 Hypothesis testing: The following are the proportions of girl births in Vienna for each month in  Girl births 1908 and 1909 (out of an average of 3900 births per month):  .4777 .4875 .4859 .4754 .4874 .4864 .4813 .4787 .4895 .4797 .4876 .4859  .4857 .4907 .5010 .4903 .4860 .4911 .4871 .4725 .4822 .4870 .4823 .4973  The data are in the folder Girls. These proportions were used by von Mises (1957) to support  a claim that that the sex ratios were less variable than would be expected under the binomial  distribution. We think von Mises was mistaken in that he did not account for the possibility that  this discrepancy could arise just by chance.  

(a) Compute the standard deviation of these proportions and compare to the standard deviation that would be expected if the sexes of babies were independently decided with a constant  probability over the 24-month period.  

**Answer:**  

Sample standard deviation:  

$s = \sqrt{\frac{\sum\limits_{i=1}^{24}(x_i - \bar{x})}{24-1}} \approx 0.0064$

mean: 

$\bar{x} = \frac{\sum\limits_{i=1}^{24}}{24} \approx 0.4857$ 

$E(sd) = \sqrt{0.4857 * (1 - 0.4857) / 3903} \approx 0.0080$ 

diff = 0.0080 - 0.0064 = 0.0016 

(b) The observed standard deviation of the 24 proportions will not be identical to its theoretical  expectation. In this case, is this difference small enough to be explained by random variation?  Under the randomness model, the actual variance should have a distribution with expected  value equal to the theoretical variance, and proportional to a χ2 random variable with 23  degrees of freedom; see page 53. 

**Answer:**  

$df = 23, \chi_{0.025}^2 \approx 11.6886$ 

$df = 23, \chi_{0.975}^2 \approx 38.0756$ 

95% CI of sd = [$\sqrt{23 * 0.0064^2 / \chi_{0.975}^2}$, $\sqrt{23 * 0.0064^2 / \chi_{0.025}^2}$] $\approx$ [0.0050, 0.0090] 

Since the observed sd (0.0064) is in the 95% CI, we conclude that the difference is not significant at 95% level and can be explained by random variation. 

4.7 Inference from a proportion with y = 0: Out of a random sample of 50 Americans, zero report  having ever held political office. From this information, give a 95% confidence interval for the  proportion of Americans who have ever held political office. 

**Answer:**  
$\hat{p} = \frac{y + 2}{n + 4} = \frac{1}{27}$

$se = \sqrt{\hat{p}(1-\hat{p})/(n + 4)} = \sqrt{\frac{1}{27}(1 - \frac{1}{27}) / 54} \approx 0.026$

$[\hat{p} \pm 2se] = [\frac{1}{27} \pm 2 * 0.026] \approx [-0.015, 0.089]$

So the 95% confidence interval is [0%, 8.9%].

4.9 Inference for a probability: A multiple-choice test item has four options. Assume that a student  taking this question either knows the answer or does a pure guess. A random sample of 100 students take the item, and 60% get it correct. Give an estimate and 95% confidence interval for the percentage in the population who know the answer. 

**Answer:**  
P(Correct|Guess) = 0.25   

P(Correct) = 0.6 

P(Correct|Know) = 1 

P(Know|Correct) = P(Correct|Know)P(Know) / P(Correct) 

P(Guess|Correct) = P(Correct|Guess)P(Guess) / P(Correct) 

P(Know|Correct) + P(Guess|Correct) = 1 

1 * P(Know) / 0.6 + 0.25 * (1 - P(Know)) / 0.6 = 1

P(Know) = (0.6 - 0.25) / 0.75 $\approx$ 0.467 

95% CI = [P(Know) $\pm$ 2 * $\sqrt{0.467 * (1 - 0.467) / 100} \approx$  [0.467 $\pm$ 0.1] = [0.367, 0.567] 

So the estimate is 47% and 95% CI is [37%, 57%].

Or to be more precise,

P(Know) = (P(Correct) - 0.25) / 0.75  

Var(P(Know)) = Var((P(Correct) - 0.25) / 0.75) = (1 / 0.75)^2^Var(P(Correct)) $\approx$ 1.778 * Var(P(Correct)) 

95% CI = [P(Know) $\pm$ 2 * $\sqrt{1.778 * 0.6 * (1 - 0.6) / 100} \approx$  [0.467 $\pm$ 2 * 0.065] = [0.337, 0.597]



