---
title: "lab_data_viz"
author: "Chenfeng Hao"
date: "9/12/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(readr)
```

## Data Viz Lab

Load data from csv file

```{r tidy=TRUE}
nc <-
  read.csv(
    "https://docs.google.com/spreadsheets/d/e/2PACX-1vTm2WZwNBoQdZhMgot7urbtu8eG7tzAq-60ZJsQ_nupykCAcW0OXebVpHksPWyR4x8xJTVQ8KAulAFS/pub?gid=202410847&single=true&output=csv"
  )
```

Take a glimpse

```{r}
glimpse(nc)
```

Question

1. What type of variable is R considering the variable habit to be? What variable type is visits? (answer
with text)

habit, characters; visits, integers.

### Scatterplot 

```{r}
ggplot(data = nc, aes(x = weeks, y = weight)) +
  geom_point()
```

Add a new labels layer  

```{r}
ggplot(data = nc, aes(x = weeks, y = weight)) +
  geom_point() +
  labs(x = "Length of pregnancy (weeks)", y = "Birth weight of baby (lbs)",
       title = "Relationship between pregnancy duration and newborn weight")
```


Questions

1. Is there a positive or negative relationship between these variables? (text only to answer)  

There is an overall positive relationship between the length of pregnancy and birth weight of babies.  

2. Make a graph showing weeks again on the x axis and the variable gained on the y axis (the amount of
weight a mother gained during pregnancy). Include axis labels with measurement units, and a title.
(code only to answer)

```{r results='hide'}
ggplot(data = nc, aes(x = weeks, y = gained)) +
  geom_point() +
  labs(x = "Length of pregnancy (weeks)", y = "Gained weight of baby (lbs)",
       title = "Relationship between pregnancy duration and newborn gained weight")
```

3. Study the code below, and the resulting graphical output. Note that I added a new argument of color
= premie inside the aesthetic mapping. The variable premie indicates whether a birth was early
(premie) or went full term. Please answer with text:

A. What did adding the argument color = premie accomplish?

Distinguishes categories by colors.

B. How many variables are now displayed on this plot?

3, pregnancy length, maternal weight gain, premie or full term. 

C. What appears to (roughly) be the pregnancy length cutoff for classifying a newborn as a “premie” versus a “full term”.

Between 36 and 37 weeks.  

```{r}
ggplot(data = nc, aes(x = weeks, y = gained, color = premie)) +
  geom_point() +
  labs(x = "Pregnancy length (wks)", y = "Maternal weight gain (lbs)")
```

Question 

1. Make a new scatter plot that shows a mothers age on the x axis (variable called mage) and birth weight of newborns on the y axis (weight). Color the points on the plot based on the gender of the resulting baby (variable called gender). Does there appear to be any strong relationship between a mother’s age and the weight of her newborn? (code and text to answer)

```{r}
ggplot(data = nc, aes(x = mage, y = weight, color = gender)) +
  geom_point() +
  labs(x = "Mother Age", y = "Newborn Birth Weight") +
  geom_smooth(method = "lm")
```

There does not appear to be any strong relationship between a mother’s age and the weight of her newborn.  

### Histogram

```{r}
ggplot(data = nc, aes(x = weeks)) +
  geom_histogram()
```


```{r}
ggplot(data = nc, aes(x = weeks)) +
  geom_histogram(binwidth = 1,
                 color = "white",
                 fill = "steelblue")
```

Questions

1.  Inspect the histogram of the weeks variable. Answer each of the following with text.

A. The y axis is labeled count. What is specifically being counted in this case? Hint: think about
what each case is in this data set.

The number of pregnancies of a particular duration, as shown on the x axis.

B. What appears to be roughly the average length of pregnancies in weeks?

The average length of pregnancies seems to be close to 39 weeks.

Average length of pregnancies in weeks

```{r}
weeks <- nc$weeks
mean(weeks)
```

C. If we changed the binwidth to 100, how many bins would there be? Roughly how many cases would
be in each bin?

If we changed the binwidth to 100, there would be only 1 bin and all cases would be in the bin, since the max number of weeks in the data is 45.

```{r}
max(weeks)

ggplot(data = nc, aes(x = weeks)) +
  geom_histogram(binwidth = 100,
                 color = "white",
                 fill = "steelblue")
```


2. Make a histogram of the birth weight of newborns (which is in lbs), including a title and axis labels.
(code only to answer)


```{r}
ggplot(data = nc, aes(x = weight)) +
  geom_histogram(binwidth = 1,
                 color = "white",
                 fill = "steelblue") +
  labs(x = "Birth weight of newborns (lbs)", y = "Number of newborns", title = "histogram of the birth weight of newborns")
```


#### Faceting


```{r}
ggplot(data = nc, aes(x = weeks)) +
  geom_histogram(binwidth = 1,
                 color = "white",
                 fill = "steelblue") +
  facet_wrap( ~ mature, ncol = 1)
```

Question

1.Make a histogram of newborn birth weight split by gender of the child. Set the binwidth to 0.5. Which gender appears to have a slightly larger average birth weight? (code and text to answer)

```{r}
ggplot(data = nc, aes(x = weight)) +
  geom_histogram(binwidth = 0.5,
                 color = "white",
                 fill = "steelblue") +
  facet_wrap(~ gender, ncol = 1)
```

The group of males seems to have a slightly larger average birth weight.

```{r}
male_weight <- filter(nc, gender == "male")$weight
mean(male_weight)

female_weight <- filter(nc, gender == "female")$weight
mean(female_weight)
```


### Boxplot

```{r}
ggplot(data = nc, aes(x = gender, y = gained)) +
  geom_boxplot(fill = "sienna")
```

Questions

8. Make a boxplot of the weight gained by moms, split by the maturity status of the mothers (mature). Include axis labels and a title on your plot. Is the median weight gain during pregnancy larger for younger or older moms? (text and code)

```{r}
ggplot(data = nc, aes(x = mature, y = gained)) +
  geom_boxplot(fill = "sienna") +
  labs(title = "Boxplot of the weight gained by the maturity status of the mothers")
```

The median weight gain during pregnancy is larger for younger moms? 

9. Make a boxplot of pregnancy duration in weeks by smoking habit. Is the duration of pregnancy more variable for smokers or non-smokers? (i.e. which group has the greater spread for the variable weeks?).(code and text to answer) 

```{r}
ggplot(data = nc, aes(x = habit, y = weeks)) +
  geom_boxplot(fill = "sienna") +
  labs(title = "Boxplot of pregnancy duration in weeks by smoking habit")
```

The duration of pregnancy is more variable for smokers. 

Questions

1. Using a data visualization, visually assess: Is the variable for father’s age (fage) symmetrical, or does
it have a skew?

```{r}
ggplot(data = nc, aes(x = fage)) +
  geom_histogram(aes(y = ..density..),
                 binwidth = 1,
                 color = "white",
                 fill = "steelblue") +
  geom_density(col="red")
```

The variable for father’s age (fage) is skewed to the right. 

2. Using a data visualization, visually assess: (in this sample) is the median birth weight of babies greater
for white or non-white mothers (variable called whitemom)? 

```{r}
ggplot(data = nc, aes(x = whitemom, y = weight)) +
  geom_boxplot(fill = "sienna")
```

The median birth weight of babies is greater for white mothers.   

3. Using a data visualization, visually assess: (in this sample) as a mother’s age (mage) increases, does the
duration of pregnancy (weeks) appear to decrease?

```{r}
ggplot(data = nc, aes(x = mage, y = weeks)) +
  geom_point() +
  labs(x = "Mother's age", y = "Duration of pregnancy (weeks)",
       title = "Relationship between pregnancy duration and mother's age") +
  geom_smooth(method='lm')


```

In this sample, as a mother’s age (mage) increases, the duration of pregnancy (weeks) does not appear to decrease much.  
